Stack Safety is a Security Property

### 2021-04-16 tex/main.tex

\definition
Let the protected set of a context, \(\prot(\domMap,\rts)\), be the
set of components \(\component\) such that \(\domMap ~ \component = \sealed{\depth}\)
for any depth \(\depth\).

\definition
A system enjoys {\em stepwise stack integrity} if, for any reachable state
\(\mpcstatename\) such that \(\mpcstatename \stepstocon \mpcstatename'\),
and any component \(\component \in \prot(\domMap,\rts)\),
\(\pi_\mach(\mpcstatename') ~ \component = \pi_\mach(\mpcstatename) ~ \component\).

@@

\definition
For each trace \(\MPCT \in \MPCT^*\) with \(\head(\MPCT) = \underscore,\underscore,(\underscore,\rts)\), the
{\em break condition} of \(\MPCT\) is a predicate on
states \(\brk(\MPCT) = \neg f_{|\rts|}\). It follows from our definitions
that this holds on \(\MPCT\)'s final state, {\em unless it terminated prematurely due to a failstop}.

@@

\definition
For any call trace \(\MPCT \in \MPCT^*\) with \(\head(\MPCT) = \mach,\pol,\context\),
let \(\components = \secret(\context)\) and \(\nach\) be any variant
\(\mach \approx_\components \nach\).
A system enjoys \emph{stepwise stack confidentiality} if,
for \(\NPCT\) such that \(\nach,\pol,\domMap \hookrightarrow \NPCT | \brk(\MPCT)\),
\(\MPCT \doteq \NPCT\) holds.

@@

In other words, stepwise stack confidentiality states that if we vary every component of the stack,
the resulting trace (starting in the same policy state) steps in lockstep with the
original trace until both have reached the break condition.

Note that \(\MPCT\) may contain nested calls; these must also respect the secrecy
of the components in \(\components\), and in addition their subtraces are also in
\(\MPCT^*\), and therefore confidentiality must hold on them separately. \(\NPCT\)
will contain these calls and returns as well, so its return stack pushes and pops
in lockstep until it breaks at the final return.

\subsection{Control Flow and Return-to-Caller}
\label{sec:RTC}

\BCP{We agreed that most or all of this belongs in Section 4...}Together,
integrity and confidentiality protect a caller's data until its callee
returns to it. However, it is possible that an intermediate caller is not
returned {\em to}, but rather returned {\em past}, to some ancestor caller!
For example, function $A$ calls function $B$, which in turn calls $C$, which
returns directly to $A$ by resetting the stack pointer to $A$'s stack frame
and branching to the address after the call from $A$ to $B$.  Strictly
speaking, this is not an information-flow violation, since neither $A$'s nor
$B$'s stack frames are improperly read or written by $C$.  But it is still
arguably unsafe in general.  For example, suppose that $B$'s continuation
after the call to $C$ takes $C$'s result, doubles it, and returns this to
$A$.  If $A$'s correct behavior after the call to $B$ returns depends on the
result it receives being a even number---i.e., $A$ is trusting that $B$'s
continuation will execute after $C$ returns, perhaps because $A$ and $B$ are
written by the same person while $C$ is an untrusted library---then $C$ can
cause it to behave incorrectly, possibly even insecurely, by returning an
odd result directly to $A$.
%
Thus, a formal definition of safe use of the stack should include the
condition that calls and returns are {\em well bracketed}, e.g., $C$ should
only return to $A$.

\definition
A system enjoys {\em return-to-caller} when, for every pair of reachable states
\(\mpcstatename = (\underscore,\underscore,(\underscore,\rts))\) and
\(\mpcstatename' = (\underscore,\underscore,(\underscore,\rts'))\) such that
\(\mpcstatename \stepsto \mpcstatename'\), if \(|\rts'| < |\rts|\), then
\(|\rts'| = |\rts|\)

\bcp{we probably want to remark that this notion is pretty
much the same as in StkTokens, just formalized a little differently.}

>>>>

\definition
Let the protected set of a context, \(\prot(\domMap,\rts)\), be the
set of components \(\component\) such that \(\domMap ~ \component = \sealed{\depth}\)
for some depth \(\depth\).

\definition
A system enjoys \textit{\textbf{stepwise stack integrity}} if, for any reachable state
\(\mpcstatename\) such that \(\mpcstatename \stepstocon \mpcstatename'\),
and any component \(\component \in \prot(\domMap,\rts)\),
\(\pi_\mach(\mpcstatename') ~ \component = \pi_\mach(\mpcstatename) ~ \component\).
\apt{but now $dm$ and $rts$ are unbound}.

@@

\definition
For each trace \(\MPCT \in \MPCT^*\) with \(\head(\MPCT) = \underscore,\underscore,(\underscore,\rts)\), the
{\em break condition} of \(\MPCT\) is a predicate on
states defined as \(\brk(\MPCT) = \neg f_{|\rts|}(\MPCT)\). It follows from our definitions
that this holds on \(\MPCT\)'s final state, {\em unless it terminated prematurely due to a failstop}.

@@

\definition
For any call trace \(\MPCT \in \MPCT^*\) with \(\head(\MPCT) = \mach,\pol,\context\),
let \(\components = \mathit{secret}(\context)\) and \(\nach\) be any variant
such that \(\mach \approx_\components \nach\).
A system enjoys \textbf{\textit{stepwise stack confidentiality}} if,
for all \(\NPCT\) such that \(\nach,\pol,\domMap \hookrightarrow \NPCT | \brk(\MPCT)\),
\(\MPCT \doteq \NPCT\) holds.\apt{I'm worried about quantifier order here. Surely $n$ needs to be
universally quantified, and probably also \(\MPCT\) (?) }

@@

In other words, stepwise stack confidentiality states that if we vary every component of the stack,
the resulting trace (starting in the same policy state) steps in lockstep with the
original trace until both have reached the break condition.
%
Note that \(\MPCT\) may contain nested calls; these must also respect the secrecy
of the components in \(\components\), and in addition their subtraces are also in
\(\MPCT^*\), and therefore confidentiality must hold on them separately. \(\NPCT\)
will contain these calls and returns as well, so its return stack pushes and pops
in lockstep until it breaks at the final return.

Finally, we formalize a version of WBCF, which capture the same informal idea
as that of \citet{Skorstengaard+19b} using our notion of contexts.
\definition
A system enjoys \textit{\textbf{well-bracketed control flow}} when, for every pair of reachable states
\(\mpcstatename = (\underscore,\underscore,(\underscore,\rts))\) and
\(\mpcstatename' = (\underscore,\underscore,(\underscore,\rts'))\) such that
\(\mpcstatename \stepsto \mpcstatename'\), if \(|\rts'| < |\rts|\), then
\(|\rts'| = |\rts|\)
